---
title: "Target Markdown"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Target Markdown is a powerful R Markdown interface for reproducible analysis pipelines, and the chapter at https://books.ropensci.org/targets/markdown.html walks through it in detail. This R Markdown report the example from the chapter. Try it out in both interactive and non-interactive modes, either by running the code chunks in different ways or setting the `tar_interactive` chunk option.

# Packages

The example requires several R packages, and `targets` must be version 0.5.0.9000 or above. 

```{r, eval = FALSE}
install.packages(c("biglm", "dplyr", "ggplot2", "readr", "targets", "tidyr"))
```

# Setup

If you are using old versions of `targets` (<= 0.7.0) and/or `knitr` (<= 1.33), you will need to load the `targets` package in the R Markdown document in order for Target Markdown code chunks to work.

```{r}
library(targets)
```

Near the top of the document, you may also wish to remove the `_targets_r` directory previously written by non-interactive runs of the report. Otherwise, your pipeline may contain superfluous targets.

```{r}
library(tidyverse)
library(usethis)
library(here)
library(janitor)
# source custom functions in R  folder
invisible(purrr::map(dir(here("R/"),full.names = T),~source(.x,verbose = FALSE,echo = F)))
tar_unscript()
```

# Globals

We first define some global options/functions common to all targets. The function below plots a histogram of ozone concentrations, and our histogram target will need it.

```{targets example-globals, tar_globals = TRUE}
# options(tidyverse.quiet = TRUE)
# tar_option_set(packages = c("tidyverse"))


root_dir <- Sys.getenv("CC_RB_LF_SCH_MONTHLY")
data_dir <- glue::glue("{root_dir}/ETH/data_raw/")

```

# Targets

Our first target borrows the `airquality` dataset built into base R.

```{targets raw-data}
tar_target(post_201905_RB,
           compile_tab(folder_path = data_dir,which_tabs = "RB_rx")
           )
```

Our next targets preprocess the data, make a histogram, and fit a model.

```{targets downstream-targets}
list(
  tar_target(data, raw_data %>% filter(!is.na(Ozone))),
  tar_target(hist, create_plot(data))
)
```

```{r}
tar_make()
```

# Output

You can retrieve results from the `_targets/` data store using `tar_read()` or `tar_load()`.

```{r, message = FALSE}
library(biglm)
tar_read(fit)
```

```{r}
tar_read(hist)
```

The `targets` dependency graph helps your readers understand the steps of your pipeline at a high level.

```{r}
tar_visnetwork()
```

At this point, you can go back and run `{targets}` chunks in interactive mode without interfering with the code or data of the non-interactive pipeline.
