---
title: "cleaning_raw_data_catalogue"
author: "Zack Arno"
date: '2022-10-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(usethis)
library(here)
library(janitor)
# source custom functions in R  folder
invisible(purrr::map(dir(here("R/"),full.names = T),~source(.x,verbose = FALSE,echo = F)))
```

## Intro

The purpose of this notebook is to document the data cleaning process. This documents walks through the data cleaning process and function generation. As concepts are turned into generic functions they are  added to the `R` directory. The actual data cleaning will take place with just a few functions chained together in a separate notebook, but this will serve as a good reference to how and which analytic data cleaning decisions were made

Here we outline several cleaning steps:

1. Compile the raw data catalog
2. Clean raw data to make clean data catalogue
  a. Read/extract/compile "relevant" tables from raw data
  b. Harmonize administrative names to master list
  c. Select relevant/critical variables from each table

## Compile the raw data catalogue

This step has been completed. The functions to do this are in the `R/` folder. I will either add the code documentation here or create a separate notbook for this process (*TODO*).

## Name Harmonization

Perhaps the most cumbersome step of the cleaning will be to harmonize all the administrative names used during data collection to a master list. I recommend using the administrative boundaries files endorsed by OCHA which can be downloaded [here](https://data.humdata.org/dataset/cod-ab-eth).

This step requires feedback from other team members and should be considered the **main initial data cleaning priority**. Therefore, I have set up a pipeline and created some tools to facilitate the cleaning of names.

First we have some convenience functions to compile the data sets and extract admin names. Note you will need to create an `system environment variable` that holds the entire path to the directory  `Country Monthly Reports` directory of the onedrive. Once this environment is stored you can run the following chunk to load all the "RB_rx" tabs. By harmonizing all the names in this tab, we will essentially be creating a admin name dictionary/lookup table that can be used to reclassify admin names in other tabs.


```{r load_data}
root_dir <- Sys.getenv("CC_RB_LF_SCH_MONTHLY")
data_dir <- glue::glue("{root_dir}/ETH/data_raw/")

dataset_list <- compile_tab(folder_path = data_dir,which_tabs = "RB_rx")
```

Next we will load and compile all the admin columns from the raw data catalog This is done with the `load_admins_from_raw` function. **Note:** The final data pipeline will likely change, but these steps are used to create our administrative name harmonization database/lookup.



```{r load_admins}
# load admin columns
admins_raw <- dataset_list |>
  purrr::map_dfr(~load_admins_from_raw(.x))

# load ocha endoresed admin boundaries 
master_adm<- read_rds(here("data/eth_adm_master"))
```

We want to sanitize admin names (i.e tolower case, trimws, remove characters/symbols), but first we need to fix a data entry/quality issue. When there was just one `admin 3` being reported in a specific `admin 2` the staff entering the data often wrote `1` into `admin 3` and in the `admin 2` column they wrote the `admin 2 - admin 3`. Therefore we will fix that here - we can turn this into a function later - for now just run this chunk


```{r fix_entry_error1}
# before sanitzing lets set adm 3 properly (when admin 2/3 sometimes combined in adm2)
admins_c1 <- admins_raw |> 
  separate(col = adm2_name,into = c("adm2_rev","adm3_rev"),sep = "-") |> 
  mutate(adm3_f= if_else(!is.na(adm3_rev),adm3_rev, adm3_name)) |> 
  filter(!adm3_f %in% as.character(c(1:1000))) |> 
  select(adm1_name, adm2_name = adm2_rev, adm3_name= adm3_f)
```

### Fix Admin 1

1. sanitize admin 1
2. check admin1 against master 
3. fix admin 1
4. check fixed admin 1 against master (repeat steps 3 & 4 until clean)
```{r fix_admin1}
# step 1 - preliminary clean of admin 1
admins_c1_san<- admins_c1 |> 
  sanitize_admins()

check_admin1(df = admins_c1,df_master = master_adm)


# with these changes it should be fixed.
cleaned_adm1_a <- admins_c1_san |> 
  mutate(
    adm1_name = case_when(
      adm1_name == "ben_gumuz"~"benishangul_gumz",
      adm1_name == "gambella"~"gambela" ,
      adm1_name == "snnpr"~"snnp", #snnp used to match HDX file
      adm1_name == "swe"~"south_west_ethiopia",
      TRUE~ adm1_name
    )
  )
    
    
check_admin1(df = cleaned_adm1_a,df_master = master_adm)
```

### Fix Admin 2

This follows the same procedure as above, but instead you can use the `check_admin2()` function.

three types of data entry errors have been noted:

1. incorrect/misspelled admin 2
2. incorrect/misspelled admin 1
3. incorrect/misspelled admin 1 & 2 

I frequently pull up the master list and sort by admin 2 to find the correct reclassification values. 
```{r fix_admin2}
check_admin2(cleaned_adm1_a,df_master = master_adm)

cleaned_adm2_a <- cleaned_adm1_a |> 
  mutate(
    # TYPE 2: INCORRECT ADMIN 1 - FIX ADMIN 1
    adm1_name= case_when(
      adm1_name == "snnp" & adm2_name %in%c("kaffa","kefa")~"south_west_ethiopia", 
      adm1_name == "snnp" & adm2_name %in%c("dawuro")~"south_west_ethiopia", 
      adm1_name == "snnp" & adm2_name %in%c("bench_sheko","bench_maji", "sheka")~"south_west_ethiopia",
      adm1_name == "snnp" & adm2_name %in%c("west_omo")~"south_west_ethiopia",
      TRUE~adm1_name
    ),
    # TYPE 1:  INCORRECT ADMIN 2 - FIX ADMIN 2
     adm2_name = case_when(
      adm2_name == "illubabor"~"ilu_aba_bora" ,
      adm2_name == "agnuwa"~"agnewak" ,
      adm2_name == "west_gojjam"~"west_gojam" ,
      adm1_name == "south_west_ethiopia" & adm2_name == "kaffa"~"kefa" ,
      adm1_name == "oromia" & adm2_name == "north_shoa"~"north_shewa_or" ,
      adm1_name == "amhara" & adm2_name == "north_shoa"~"north_shewa_am" ,
      adm1_name == "oromia" & adm2_name == "sw_shoa"~"south_west_shewa" ,
      adm1_name == "gambela" & adm2_name == "mezheng"~"majang" ,
      adm1_name == "amhara" & adm2_name == "south_wollo"~"south_wello" ,
      adm1_name == "gambela" & adm2_name == "itang_sp_w"~"itang_special_woreda" ,
      adm1_name == "oromia" & adm2_name == "e_arsi"~"arsi" ,
      adm1_name %in%c("snnp", "snnpr") & adm2_name =="kaffa"~"kefa",
      adm1_name == "benishangul_gumz" & adm2_name == "dam_workers_metekel"~"metekel" ,
      adm1_name == "gambela" & adm2_name == "refugees_7_camps"~"refugees" ,
      adm1_name == "gambela" & adm2_name == "refugees_7_vamps"~"refugees" ,
      adm1_name == "gambela" & adm2_name == "refugees_gambella"~"refugees" ,
      adm2_name == "gamogofa" ~ "gofa",
      adm2_name == "west_omo" ~ "mirab_omo",
      adm2_name == "west_om_b_149_o" ~ "mirab_omo",
      adm2_name == "bench_maji" ~ "bench_sheko",

      TRUE ~ adm2_name
    )
    
  ) 


master_adm |>
  distinct(adm1_en,adm2_en, adm3_en) |>
  arrange(
    adm3_en
    )  
   
check_admin2(cleaned_adm2_a,df_master = master_adm)  

```


# Fix admin 3

```{r}
## this will be handy for maually comparing to master
master_adm_adj <-  master_adm |>
    sanitize_admins() |>
    mutate(
      adm_1_2_3=  glue::glue("{adm1_en}-{adm2_en}-{adm3_en}")
    )

cleaned_adm3_a <- cleaned_adm2_a |> 
  mutate(
    # lets separate admins with "-" to avoid confusion
    adm_1_2_3=  glue::glue("{adm1_name}-{adm2_name}-{adm3_name}"),
    
    # SCENARIO 1: adm1 mixed up - uncomment and fill if necessary
    # adm1_name= case_when()
    
    # SCENARIO 2: adm2 mixed upp
    adm2_name = case_when(
      adm_1_2_3=="amhara-central_gondar-quara"~ "west_gondar",
      adm_1_2_3 == "snnp-gamo-gezie_gofa" ~ "gofa",
      adm_1_2_3 == "snnp-gamo-melokoza" ~ "gofa",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-meneat_shasha" ~ "mirab_omo",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-meneat_goldia" ~ "mirab_omo",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-maji" ~ "mirab_omo",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-bero" ~ "mirab_omo",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-surma" ~ "mirab_omo",
      TRUE~ adm2_name
    ),
    
    adm_1_2_3=  glue::glue("{adm1_name}-{adm2_name}-{adm3_name}"),
    
    # SCENARIO 3 adm3 wrong
    adm3_name = case_when(
      adm_1_2_3=="amhara-south_gondar-e_estie"~ "east_esite",
      adm_1_2_3 %in% c("amhara-south_gondar-w_estie_andabet","amhara-south_gondar-w_estie")~ "andabet_west_esite",
      adm_1_2_3=="amhara-west_gojam-s_achefer"~ "debub_achefer",
      adm_1_2_3=="amhara-west_gojam-n_achefer"~ "semen_achefer",
      adm_1_2_3=="gambela-agnewak-dimma"~ "dima_gm",
      adm_1_2_3=="gambela-majang-goderie"~ "godere",
      adm_1_2_3=="oromia-buno_bedele-bedele"~ "bedele_zuria",
      adm_1_2_3=="oromia-buno_bedele-bedele_t"~ "bedele_town",
      adm_1_2_3=="gambela-agnewak-gambella_town"~ "gambela_town",
      adm_1_2_3=="gambela-agnewak-gambella_zuria"~ "gambela_zuria",
      adm_1_2_3 == "amhara-south_gondar-simada" ~ "semada",
      adm_1_2_3 == "benishangul_gumz-metekel-dibatie" ~ "dibate",
      adm_1_2_3 == "benishangul_gumz-metekel-NA" ~ "dam_workers",
      adm_1_2_3 == "benishangul_gumz-metekel-0" ~ "dam_workers",
      adm_1_2_3 == "benishangul_gumz-metekel-wombera" ~ "wembera",
      adm_1_2_3 == "gambela-agnewak-gambella" ~ "gambela_zuria",
      adm_1_2_3 == "gambela-itang_special_woreda-itang_sp_w" ~"itang",
      adm_1_2_3 == "gambela-majang-mengeshi" ~ "mengesh",
      adm_1_2_3 == "gambela-refugees-0" ~ "refugees",
      adm_1_2_3 == "gambela-refugees-NA" ~ "refugees",
      adm_1_2_3 == "oromia-arsi-digaluna_tijo" ~ "degeluna_tijo",
      adm_1_2_3=="south_west_ethiopia-dawuro-terch_town"~  "tercha_town",
      adm_1_2_3=="south_west_ethiopia-dawuro-tarcha_zuria"~  "tercha_zuriya",
      adm_1_2_3=="south_west_ethiopia-kefa-goba"~  "goba_sp",
      adm_1_2_3== "oromia-buno_bedele-chewaka" ~ "chwaka",
      adm_1_2_3== "oromia-buno_bedele-chora" ~ "chora_(buno_bedele)",
      adm_1_2_3 == "oromia-buno_bedele-didesa" ~ "dedesa",
      adm_1_2_3 == "oromia-east_hararge-deder_rural" ~ "deder", 
      adm_1_2_3 == "oromia-east_hararge-gursum" ~ "gursum_(or)",
      adm_1_2_3 == "oromia-east_hararge-jarso" ~ "jarso_(east_hararghe)",
      adm_1_2_3 == "oromia-east_hararge-kersa" ~ "kersa_(east_hararge)",
      adm_1_2_3 == "oromia-east_hararge-melka_belo" ~ "melka_balo",
      adm_1_2_3 == "oromia-east_hararge-meyumulke" ~ "meyu_muleke",
      adm_1_2_3 == "oromia-ilu_aba_bora-algie_sachi" ~ "alge_sachi",
      adm_1_2_3 == "oromia-ilu_aba_bora-becho" ~ "becho_(ilu_aba_bora)",
      adm_1_2_3 == "oromia-ilu_aba_bora-bure" ~ "bure_(or)",
      adm_1_2_3 == "oromia-ilu_aba_bora-halu" ~ "halu /huka",
      adm_1_2_3 == "oromia-ilu_aba_bora-hurrumu" ~ "hurumu",
      adm_1_2_3 == "oromia-ilu_aba_bora-mettu" ~ "metu_zuria",
      adm_1_2_3 == "oromia-ilu_aba_bora-mettu_town" ~ "metu_town",
      adm_1_2_3 == "oromia-ilu_aba_bora-yayo" ~ "yayu",
      adm_1_2_3 == "oromia-jimma-agaro" ~ "agaro_town",
      adm_1_2_3 == "oromia-jimma-chora_botor" ~ "chora_(jimma)",
      adm_1_2_3 == "oromia-jimma-gomma" ~ "goma",
      adm_1_2_3 == "oromia-jimma-gumey" ~ "gumay",
      adm_1_2_3 == "oromia-jimma-kersa" ~ "kersa_(jimma)",
      adm_1_2_3 == "oromia-jimma-limu_saka" ~ "limu_seka",
      adm_1_2_3 == "oromia-jimma-mana" ~ "mena_(jimma)",
      adm_1_2_3 == "oromia-jimma-mencho" ~ "mancho",
      adm_1_2_3 == "oromia-jimma-shebe_sembo" ~ "shebe_sambo",
      adm_1_2_3 == "oromia-jimma-sokoru" ~ "sekoru",
      adm_1_2_3 == "oromia-north_shewa_or-girar_jarso" ~ "gerar_jarso",
      adm_1_2_3 == "oromia-north_shewa_or-hidhabu_abote" ~ "hidabu_abote",
      adm_1_2_3 == "oromia-west_hararge-chiro_rural" ~ "chiro_zuria",
      adm_1_2_3 == "oromia-west_hararge-guba_qoricha" ~ "goba_koricha",
      adm_1_2_3 == "oromia-west_hararge-oda_bultum" ~ "kuni_/oda_bultum",
      adm_1_2_3 == "oromia-west_hararge-tulo" ~ "tulo_(or)",
      adm_1_2_3 == "snnp-gofa-gezie_gofa" ~ "gezei_gofa",
      adm_1_2_3 == "snnp-gofa-melokoza" ~ "melekoza",
      adm_1_2_3 == "oromia-west_hararge-burka_dimtu" ~ "burqua_dhintu",
      adm_1_2_3 == "snnp-south_omo-selamago" ~ "salamago",
      adm_1_2_3 == "snnp-south_omo-selamego" ~ "salamago",
      adm_1_2_3 == "snnp-south_omo-semen_ari" ~ "north_ari",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-gide_bench" ~ "gidi_bench",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-guraferda" ~ "gurafereda",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-mizan_aman" ~ "mizan_aman_town",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-shie_bench" ~ "shay_bench",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-siz_town" ~ "size_town",
      adm_1_2_3 == "south_west_ethiopia-dawuro-esera" ~ "isara",
      adm_1_2_3 == "south_west_ethiopia-dawuro-essera" ~ "isara",
      adm_1_2_3 == "south_west_ethiopia-dawuro-gena_bossa" ~ "gena",
      adm_1_2_3 == "south_west_ethiopia-dawuro-kechi" ~ "kachi",
      adm_1_2_3 == "south_west_ethiopia-dawuro-mermanisa" ~ "mari_mansa",
      adm_1_2_3 == "south_west_ethiopia-dawuro-tercha_city" ~ "tercha_town",
      adm_1_2_3 == "south_west_ethiopia-dawuro-zaba_gazo" ~ "zabagazo",
      adm_1_2_3 == "south_west_ethiopia-kefa-adiyo" ~ "adiyio",
      adm_1_2_3 == "south_west_ethiopia-kefa-bitta" ~ "bita",
      adm_1_2_3 == "south_west_ethiopia-kefa-bonga" ~ "bonga_town",
      adm_1_2_3 == "south_west_ethiopia-kefa-shishoende" ~ "shisho_ande",
      adm_1_2_3 == "south_west_ethiopia-kefa-tello" ~ "tullo",
      adm_1_2_3 == "south_west_ethiopia-mirab_omo-gorigesha" ~ "gori_gesha",
      adm_1_2_3 == "south_west_ethiopia-mirab_omo-meneat_goldia" ~ "menit_goldiye",
      adm_1_2_3 == "south_west_ethiopia-mirab_omo-meneat_shasha" ~ "menit_shasha",
      adm_1_2_3 == "south_west_ethiopia-sheka-andracha" ~ "anderacha",
      adm_1_2_3 == "south_west_ethiopia-sheka-masha_woreda" ~ "masha",
      adm_1_2_3 == "south_west_ethiopia-sheka-teppi_town" ~ "tepi",
      adm_1_2_3 == "south_west_ethiopia-sheka-yeki_woreda" ~ "yeki",
      
      TRUE~ adm3_name
    )
    
  )

check_admin3(df = cleaned_adm3_a,df_master = master_adm)


master_adm_adj |> 
  as_tibble() |> 
  arrange(adm3_en) |>
  print(n=2000) |> 
  filter(
    ### you can filter the concatenated 1_2_3
    str_detect(adm_1_2_3,"gira"),
    
    
    ### if that doesn't work comment it out and
    ### try some pattern matching on other admins
    ###**hint:** "^" = starts with
    ###**        "$" = ends with
   
    # str_detect(adm3_en, "ded")
    # str_detect(adm2_en,"north_shewa_am")
    
    )  |> 
  print(n=1000)


```

## Putting it all together

All the previoous steps have been functionalized and can be implemented like this:

```{r}

cleaned_phase1 <- dataset_list |> 
  purrr::map2(.y = names(dataset_list),
  ~ {
   .x |> 
    dplyr::mutate(
      file_name = .y
    ) |> 
      parse_top_table() |> 
      clean_names_and_admins(data_format = "current")
  }
)
```


## Pre May 2019 RB Rx

Now I am building out the functions to clean the old format data - trying to do this by just adding arguments to the already made functions (for the most part).

```{r}

# compile the ACTIVE TX UTG2 tabs which were the old format for the RB Rx data
pre_201905_rbrx<- compile_tab(folder_path = data_dir, which_tabs = "Active TX UTG2",skip = 2 )

# old format data includes only adm2 -- therefore first step to harmonize the names
# is to use pre-existing reclassifications based on just adm2 - thats what `clean_adm2` doesin this case
# however it wont be perfect and will require we add some more reclassificaitons

pre_201905_rbrx_clean1 <- pre_201905_rbrx |> 
  purrr::map( ~ {
    .x |>
      parse_top_table() |> 
      janitor::clean_names() |> 
      standardize_admin_names(data_format = "old") |> 
       drop_summary_rows(data_format = "old") |>
      sanitize_admins() |> 
      clean_adm2(data_format = "old")
      
  }
  )
```


Here we check all of the pre may 2019 adm2 names used against the master list


```{r}

# make unique list of all pre may 2019 adm2 names (they have been cleaned as much as they could based on previous `case_when`)
 pre_201905_adm2_unique <- pre_201905_rbrx_clean1 |> 
  purrr::map_dfr(~.x |> 
                   select(adm2_name)
 ) |> 
   distinct() 

pre_201905_rbrx_clean1$`201601_ETH_CCRBLFSCHI_Monthly.xlsx` |> colnames()
# now here are the problems - if no adm1_en is listed that means adm2_name didn't match anything in the master.

 pre_201905_adm2_unique |> 
   left_join(
     master_adm_adj |> 
       select(adm1_en, adm2_en) |> 
       distinct(),
     by= c("adm2_name"="adm2_en")
     ) |> 
   filter(is.na(adm1_en)) |> 
   print(n=40)

```




While waiting on input to ``clean_adm2()` i can inspect old col names and fix based on new.
```{r,eval =T}

rename_lookup<- readxl::read_xlsx(path = "colname_harmonization_lookup.xlsx") 

pre_201905_rbrx_clean1 |> 
  purrr::map2(.y= names(pre_201905_rbrx_clean1), ~{
    print(.y)
               rename_cols_with_lookup(df = .x,lookup = rename_lookup,lookup_fixed = F)
             
  }
  )
    



```



## Select relevant/critical variables

The above steps have been combined into generic functions (in the R directory). A satisfactory amount of names have been cleaned and it is time to move onto the next phase where we start to compile the data.

In the chunk below all of these steps developed above are run on all of the data sets loaded from the raw data catalog. The results is a list containing data.frames




Now we need to figure out which columns are important to keep for the compiled data set. Let's make a frequency table of columns names.

When you print the table below you can see that the column names look pretty clean. For almost all columns there are 37 entries (i.e the column was present in all data sets). `x28` and `x8` are most likely junk that can be disregarded. The only columns that do not show up in all 37 data sets are `total_popn_census` & `total_popn_projected`. However there frequencies combined add up to 36 which is suspiciously close to 37. It's likely that there was a column name switch at some point


```{r}
column_name_freq_table <- cleaned_phase1 |> 
  purrr::map_dfr(
    ~data.frame(col_names= colnames(.x))  
  ) |> 
  count(col_names)

column_name_freq_table |> 
    write.table("clipboard", sep="\t", row.names=FALSE)
```


### Inputs needed
To clear up this issue set the parameters below
```{r}
# choose the column name to go with for total pop from here on out-- by default i chose the one which showed more records
totalpop_col_name_to_use <- "total_popn_projected" # choose the colname to use if

# here i have pasted all the column names - please comment out columns that are not relevant for our work at the moment (we can always add back in later). The less columns the simpler the work will be going forward.

critical_cols_to_keep <-  c("active_villages_for_the_year",
                            "adm1_name",
                            "adm2_name",
                            "adm3_name",
                            "file_name",
                            "percent_active_villages_treated_for_round_1",
                            "percent_active_villages_treated_for_round_2",
                            "percent_popn_treated_round_1",
                            "percent_popn_treated_round_2",
                            "percent_utg_2_treated_all_rounds",
                            "percent_utg_treated_round_1",
                            "popn_treated_cumulative_all_rounds",
                            "popn_treated_during_current_month",
                            "popn_treated_round1",
                            "popn_treated_round2",
                            "siz_cross_border_yes_no",
                            "total_popn_census",     # comment out name we do not need to use   
                            "total_popn_projected" , # " " 
                            "utg_2_treatment_target_for_the_whole_year",
                            "utg_treatment_target_for_each_round",
                            "villages_treated_during_current_month",
                            "villages_treated_round_1",
                            "villages_treated_round_2")
```

## Let's compile

- replace column name
- remove junk cols
```{r}

pop_col_options <- c("total_popn_projected","total_popn_census")
replace_this <-  setdiff(pop_col_options,totalpop_col_name_to_use)
cleaned_compiled_phase_1a <-cleaned_phase1 |> 
  purrr::map_dfr(
    ~{
      x_temp <-  .x
      if(replace_this %in% colnames(.x)){
       x_temp <-  .x |> 
          rename(!!totalpop_col_name_to_use:=replace_this)
          
      }
      x_temp |>
        dplyr::select(-dplyr::any_of(matches("^x\\d*")))
    }
  )

```

Now we can investigate as one data set!

```{r}
cleaned_compiled_phase_1b <- cleaned_compiled_phase_1a |> 
  mutate(
    year=str_sub(string = file_name,start = 1,end = 4) |> as.numeric(),
    month=str_sub(string = file_name,start = 5,end = 6) |> as.numeric(),
    date= lubridate::ymd(glue::glue("{year}-{month}-01"))
  ) 


cleaned_compiled_phase_1b |> 
  group_by(date, adm1_name,adm2_name,adm3_name) |> 
  count() |> 
  arrange(adm1_name,adm2_name, adm3_name) |> 
  print(n=1500)
  
cleaned_compiled_phase_1b |> 
  group_by(date) |> 
  count()
```
