---
title: "Data Quality Visuals"
output: html_document
date: "2023-01-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,include = F)
```


```{r cars}
library(targets)
library(tidyverse)
tar_load_everything()
lapply(list.files("R", full.names = TRUE, recursive = TRUE), source)
```

```{r}



# RB_post201905_adm2 %>%
RB_pre_post_compiled %>%   
  left_join(eth_master_adm %>%
              bind_rows(cc_operational()) %>% 
              distinct(adm1_en,adm1_pcode,adm2_en,adm2_pcode), c("adm1_name"= "adm1_en","adm2_name"= "adm2_en")) %>% 
  select(year, month,date, adm1_name, adm1_pcode, adm2_name,adm2_pcode,everything()) %>% 
  filter(is.na(adm1_pcode)) %>% count(adm1_name,adm2_name,file_name) %>% 
  print(n=27)


RB_pre201905_adm2
RB_post201905_adm2

eth_master_adm %>% 
  filter(str_detect(adm2_en,"north")) %>% 
  count(adm1_en, adm2_en,adm3_en)

RB_post201905_adm3 %>% 
  filter(str_detect(adm2_name,"agnewak")) %>% 
  distinct(adm1_name,adm2_name,adm3_name)

eth_master_adm %>% 
  glimpse()
  group_by(adm1_en, adm1_pcode) %>% 
  group_modify(~ add_row(.x,.before=0))
```

## Overview

### yearly % treated

The current dashboard has a heatchart like this showing % treated year (cumulative treated/utg). From the heat chart it's clear that there are data quality issues.

```{r, include=T,results='asis'}
# debugonce(heat_map_gg)
b4_fixing <- heat_map_gg_cum_adm2(.dat = RB_pre_post_compiled,x = "popn_treated_during_current_month",grp_vars = c("year","adm1_name","adm2_name"),date_col = "date")+
  ggtitle("heat chart - % pop treated by month",subtitle = "before attempting to fix spill over")

ggiraph::girafe(ggobj = b4_fixing)
```

Might want to move this code later... but i'm going to make a name patch
```{r}

RB_pre_post_compiled %>% 
  filter(str_detect(adm2_name,"refugee")) %>% distinct(adm1_name,adm2_name)

RB_post201905_adm3 %>% 
  filter(str_detect(adm3_name,"refugee")) %>% distinct(adm2_name,adm3_name)

eth_master_adm %>% 
  filter(str_detect(adm2_en,"buno")) %>% distinct(adm1_en,adm2_en,adm3_en)

eth_master_adm %>% 
  filter(str_detect(adm2_en,"gamo")) %>% distinct(adm1_en,adm2_en,adm3_en)

eth_master_adm %>% 
  filter(str_detect(adm1_en,"gamb")) %>% distinct(adm1_en,adm2_en,adm3_en)
  

# RB_pre_post_compiled %>%
#   mutate(
#     adm2_name = case_when(
#       adm2_name=="north_gondar"~"west_gondar"
#       
#       )
#   )

```

I tried fixing the spillover affects based on various set rules/algorithms, but it didn't work. It seems there are too many different ways the spillover could occur and reasons that I don't understand. Therefore, I think we might have to go with a more manual approach.

The above visual is interesting and potentially useful especially as a more final visual once the data is properly cleaned however. I think it could be more useful to make similar plot just showing the monthly treated values at the different admin levels

```{r, eval=F, include=F,results='asis'}


# In order to fix:  I reset Jan, Feb, and March monthly treated numbers to 0 if they were with 5 % of the most recent treated number of the previous year (including October, November, December).

# IF you look at `ilu_aba_boar` in year 2017 you see that 1.36 M treated in June.


# *central_gondar* month 1 2020 --- that's probably the remaining 50%. should just do at admin 3 level¬

after_fixing <- heat_map_gg(.dat = RB_pre_post_compiled,
                            x = "popn_treated_during_current_month_fix",
                            grp_vars = c("year","adm1_name","adm2_name"),date_col = "date")+
    ggtitle("heat chart - % pop treated by month",subtitle = "after attempting to fix spill over")

ggiraph::girafe(ggobj = after_fixing
                # ,width_svg = 16, height_svg =5
                )
```


## monthly treatement values

Okay I've changed the aesthetics to try to make some points more clear.

Below I've plotted the population treated per admin of interest per month. `dark grey` represents months when that admin zone was missing from the data. This can be fixed -- I can add those admins with 0 pop treated for those dates, but leaving it like this for now because perhaps it could be useful for identifying operational name adjustments/changes.

### Admin 2 (zone) levle

If you hover over the differnt tiles you notice there are some columns of tiles that have no hover... these are missing files we see them here:

- "2017-08"
- "2017-09"
- "2017-11"
- "2018-05"
- "2019-03"

```{r, include=T, results="asis"}
plot_heat_monthly_adm2 <- RB_pre_post_compiled %>% 
  filter(date<"2019-05-01") %>% 
  summarise_ploting_stats(x ="popn_treated_during_current_month",
                          grp_vars = c("year","adm1_name","adm2_name"),
                          date_col = "date") %>%
  complete(nesting(adm1_name,adm2_name),date, fill = list(popn_treated_during_current_month=NA)) %>% 
  heat_map_gg(x = date, y=adm2_name,
              fill = popn_treated_during_current_month,
              facet_var = adm1_name)

ggiraph::girafe(ggobj = plot_heat_monthly_adm2
                # ,width_svg = 16, height_svg =5
                )

# RB_pre_post_compiled %>% 
#   filter(date=="2017-08-01")
```


### Admin 3 (ward) level

this will be fun... heres a gigantic chart :-)


We can see the following files are missing:

- "2019-12" , 
- "2021-02",
- "2022-06",
- "2022-07"

```{r, include=T, results="asis"}
plot_heat_monthly_adm3 <- RB_post201905_adm3 %>% 
  filter(date>="2019-05-01") %>% 
  summarise_ploting_stats(x ="popn_treated_during_current_month",
                          grp_vars = c("year","adm1_name","adm2_name","adm3_name"),
                          date_col = "date") %>%
  complete(nesting(adm1_name,adm2_name,adm3_name),date, fill = list(popn_treated_during_current_month=NA)) %>% 
  heat_map_gg(x = date, 
              y=adm3_name,
              fill = popn_treated_during_current_month,
              facet_var = adm2_name)+
  theme(
  # axis.text.x = element_text(angle = 90),  
  text = element_text(size= 15)
  # axis.title.y = element_text(size=13)
  )
  


ggiraph::girafe(ggobj = plot_heat_monthly_adm3
                ,width_svg = 16, height_svg =64
                )

```


```{r}

adm3_level_names =RB_post201905_df_ls |>
  map_dfr(
    ~.x %>% 
      janitor::clean_names() |>
    standardize_admin_names(data_format = "current")|>
    drop_summary_rows() |>
    sep_adm_2_3() |>
    sanitize_admins() %>% 
      select(adm1_name,adm2_name,adm3_name)

    
  )


adm3_level_names %>% 
  filter(str_detect(adm2_name,"refugee")) %>% 
  distinct()
    



adm2_level_names<- RB_pre201905_df_ls %>% 
    map_dfr(
    ~.x %>% 
       parse_top_table() |>
          janitor::clean_names() |>
          standardize_admin_names(data_format = "old") |>
          drop_summary_rows(data_format = "old") |>
          sanitize_admins() %>%
      select(starts_with("adm"))
          
          #clean_adm2(data_format = data_format)
    )
}
adm2_level_names %>% filter(
  str_detect(adm2_name, "ref")
  ) %>% 
  distinct()    

RB_pre_post_compiled %>% 
  filter(str_detect(adm2_name, "refugee")) %>% 
  distinct(adm1_name,adm2_name)
RB_post201905_adm3 %>% 
  filter(str_detect(adm2_name, "refugee")) %>% 
  distinct(adm1_name,adm2_name)

RB_pre201905_adm2 %>% 
  filter(str_detect(adm2_name, "refugee")) %>% 
  distinct(adm1_name,adm2_name)

````

