---
title: "cleaning_raw_data_catalogue"
author: "Zack Arno"
date: '2022-10-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(usethis)
library(here)
library(janitor)
# source custom functions in R  folder
invisible(purrr::map(dir(here("R/"),full.names = T),~source(.x,verbose = FALSE,echo = F)))
```

## Intro

The purpose of this notebook is to document the data cleaning process. Here we outline several cleaning steps:

1. Compile the raw data catalog
2. Clean raw data to make clean data catalogue
  a. Read/extract/compile "relevant" tables from raw data
  b. Harmonize administrative names to master list
  c. Select relevant/critical variables from each table

## Compile the raw data catalogue

This step has been completed. The functions to do this are in the `R/` folder. I will either add the code documentation here or create a separate notbook for this process (*TODO*).

## Name Harmonization

Perhaps the most cumbersome step of the cleaning will be to harmonize all the administrative names used during data collection to a master list. I recommend using the administrative boundaries files endorsed by OCHA which can be downloaded [here](https://data.humdata.org/dataset/cod-ab-eth).

This step requires feedback from other team members and should be considered the **main initial data cleaning priority**. Therefore, I have set up a pipeline and created some tools to facilitate the cleaning of names.

First we have some convenience functions to compile the data sets and extract admin names. Note you will need to create an `system environment variable` that holds the entire path to the directory  `Country Monthly Reports` directory of the onedrive. Once this environment is stored you can run the following chunk to load all the "RB_rx" tabs. By harmonizing all the names in this tab, we will essentially be creating a admin name dictionary/lookup table that can be used to reclassify admin names in other tabs.


```{r load_data}
root_dir <- Sys.getenv("CC_RB_LF_SCH_MONTHLY")
data_dir <- glue::glue("{root_dir}/ETH/data_raw/")

dataset_list <- compile_tab(folder_path = data_dir,which_tabs = "RB_rx")
```

Next we will load and compile all the admin columns from the raw data catalog This is done with the `load_admins_from_raw` function. **Note:** The final data pipeline will likely change, but these steps are used to create our administrative name harmonization database/lookup.



```{r load_admins}
# load admin columns
admins_raw <- dataset_list |>
  purrr::map_dfr(~load_admins_from_raw(.x))

# load ocha endoresed admin boundaries 
master_adm<- read_rds(here("data/eth_adm_master"))
```

We want to sanitize admin names (i.e tolower case, trimws, remove characters/symbols), but first we need to fix a data entry/quality issue. When there was just one `admin 3` being reported in a specific `admin 2` the staff entering the data often wrote `1` into `admin 3` and in the `admin 2` column they wrote the `admin 2 - admin 3`. Therefore we will fix that here - we can turn this into a function later - for now just run this chunk


```{r fix_entry_error1}
# before sanitzing lets set adm 3 properly (when admin 2/3 sometimes combined in adm2)
admins_c1 <- admins_raw |> 
  separate(col = adm2_name,into = c("adm2_rev","adm3_rev"),sep = "-") |> 
  mutate(adm3_f= if_else(!is.na(adm3_rev),adm3_rev, adm3_name)) |> 
  filter(!adm3_f %in% as.character(c(1:1000))) |> 
  select(adm1_name, adm2_name = adm2_rev, adm3_name= adm3_f)
```

### Fix Admin 1

1. sanitize admin 1
2. check admin1 against master 
3. fix admin 1
4. check fixed admin 1 against master (repeat steps 3 & 4 until clean)
```{r fix_admin1}
# step 1 - preliminary clean of admin 1
admins_c1_san<- admins_c1 |> 
  sanitize_admins()

check_admin1(df = admins_c1,df_master = master_adm)


# with these changes it should be fixed.
cleaned_adm1_a <- admins_c1_san |> 
  mutate(
    adm1_name = case_when(
      adm1_name == "ben_gumuz"~"benishangul_gumz",
      adm1_name == "gambella"~"gambela" ,
      adm1_name == "snnpr"~"snnp", #snnp used to match HDX file
      adm1_name == "swe"~"south_west_ethiopia",
      TRUE~ adm1_name
    )
  )
    
    
check_admin1(df = cleaned_adm1_a,df_master = master_adm)
```

### Fix Admin 2

This follows the same procedure as above, but instead you can use the `check_admin2()` function.

three types of data entry errors have been noted:

1. incorrect/misspelled admin 2
2. incorrect/misspelled admin 1
3. incorrect/misspelled admin 1 & 2 

I frequently pull up the master list and sort by admin 2 to find the correct reclassification values. 
```{r fix_admin2}
check_admin2(cleaned_adm1_a,df_master = master_adm)

cleaned_adm2_a <- cleaned_adm1_a |> 
  mutate(
    # TYPE 2: INCORRECT ADMIN 1 - FIX ADMIN 1
    adm1_name= case_when(
      adm1_name == "snnp" & adm2_name %in%c("kaffa","kefa")~"south_west_ethiopia", 
      adm1_name == "snnp" & adm2_name %in%c("dawuro")~"south_west_ethiopia", 
      adm1_name == "snnp" & adm2_name %in%c("bench_sheko","bench_maji", "sheka")~"south_west_ethiopia",
      adm1_name == "snnp" & adm2_name %in%c("west_omo")~"south_west_ethiopia",
      TRUE~adm1_name
    ),
    # TYPE 1:  INCORRECT ADMIN 2 - FIX ADMIN 2
     adm2_name = case_when(
      adm2_name == "illubabor"~"ilu_aba_bora" ,
      adm2_name == "agnuwa"~"agnewak" ,
      adm2_name == "west_gojjam"~"west_gojam" ,
      adm1_name == "south_west_ethiopia" & adm2_name == "kaffa"~"kefa" ,
      adm1_name == "oromia" & adm2_name == "north_shoa"~"north_shewa_or" ,
      adm1_name == "amhara" & adm2_name == "north_shoa"~"north_shewa_am" ,
      adm1_name == "oromia" & adm2_name == "sw_shoa"~"south_west_shewa" ,
      adm1_name == "gambela" & adm2_name == "mezheng"~"majang" ,
      adm1_name == "amhara" & adm2_name == "south_wollo"~"south_wello" ,
      adm1_name == "gambela" & adm2_name == "itang_sp_w"~"itang_special_woreda" ,
      adm1_name == "oromia" & adm2_name == "e_arsi"~"arsi" ,
      adm1_name %in%c("snnp", "snnpr") & adm2_name =="kaffa"~"kefa",
      adm1_name == "benishangul_gumz" & adm2_name == "dam_workers_metekel"~"metekel" ,
      adm1_name == "gambela" & adm2_name == "refugees_7_camps"~"refugees" ,
      adm1_name == "gambela" & adm2_name == "refugees_7_vamps"~"refugees" ,
      adm1_name == "gambela" & adm2_name == "refugees_gambella"~"refugees" ,
      adm2_name == "gamogofa" ~ "gofa",
      adm2_name == "west_omo" ~ "mirab_omo",
      adm2_name == "west_om_b_149_o" ~ "mirab_omo",
      adm2_name == "bench_maji" ~ "bench_sheko",

      TRUE ~ adm2_name
    )
    
  ) 


master_adm |>
  distinct(adm1_en,adm2_en, adm3_en) |>
  arrange(
    adm3_en
    )  
   
check_admin2(cleaned_adm2_a,df_master = master_adm)  

```


# Fix admin 3

```{r}
## this will be handy for maually comparing to master
master_adm_adj <-  master_adm |>
    sanitize_admins() |>
    mutate(
      adm_1_2_3=  glue::glue("{adm1_en}-{adm2_en}-{adm3_en}")
    )

cleaned_adm3_a <- cleaned_adm2_a |> 
  mutate(
    # lets separate admins with "-" to avoid confusion
    adm_1_2_3=  glue::glue("{adm1_name}-{adm2_name}-{adm3_name}"),
    
    # SCENARIO 1: adm1 mixed up - uncomment and fill if necessary
    # adm1_name= case_when()
    
    # SCENARIO 2: adm2 mixed upp
    adm2_name = case_when(
      adm_1_2_3=="amhara-central_gondar-quara"~ "west_gondar",
      adm_1_2_3 == "snnp-gamo-gezie_gofa" ~ "gofa",
      adm_1_2_3 == "snnp-gamo-melokoza" ~ "gofa",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-meneat_shasha" ~ "mirab_omo",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-meneat_goldia" ~ "mirab_omo",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-maji" ~ "mirab_omo",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-bero" ~ "mirab_omo",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-surma" ~ "mirab_omo",
      TRUE~ adm2_name
    ),
    
    # SCENARIO 3 adm3 wrong
    adm3_name = case_when(
      adm_1_2_3=="amhara-south_gondar-e_estie"~ "east_esite",
      adm_1_2_3 %in% c("amhara-south_gondar-w_estie_andabet","amhara-south_gondar-w_estie")~ "andabet_west_esite",
      adm_1_2_3=="amhara-west_gojam-s_achefer"~ "debub_achefer",
      adm_1_2_3=="amhara-west_gojam-n_achefer"~ "semen_achefer",
      adm_1_2_3=="gambela-agnewak-dimma"~ "dima_gm",
      adm_1_2_3=="gambela-majang-goderie"~ "godere",
      adm_1_2_3=="oromia-buno_bedele-bedele"~ "bedele_zuria",
      adm_1_2_3=="oromia-buno_bedele-bedele_t"~ "bedele_town",
      adm_1_2_3=="gambela-agnewak-gambella_town"~ "gambela_town",
      adm_1_2_3=="gambela-agnewak-gambella_zuria"~ "gambela_zuria",
      adm_1_2_3 == "amhara-south_gondar-simada" ~ "semada",
      adm_1_2_3 == "benishangul_gumz-metekel-dibatie" ~ "dibate",
      adm_1_2_3 == "benishangul_gumz-metekel-NA" ~ "dam_workers",
      adm_1_2_3 == "benishangul_gumz-metekel-0" ~ "dam_workers",
      adm_1_2_3 == "benishangul_gumz-metekel-wombera" ~ "wembera",
      adm_1_2_3 == "gambela-agnewak-gambella" ~ "gambela_zuria",
      adm_1_2_3 == "gambela-itang_special_woreda-itang_sp_w" ~"itang",
      adm_1_2_3 == "gambela-majang-mengeshi" ~ "mengesh",
      adm_1_2_3 == "gambela-refugees-0" ~ "refugees",
      adm_1_2_3 == "gambela-refugees-NA" ~ "refugees",
      adm_1_2_3 == "oromia-arsi-digaluna_tijo" ~ "degeluna_tijo",
      adm_1_2_3=="south_west_ethiopia-dawuro-terch_town"~  "tercha_town",
      adm_1_2_3=="south_west_ethiopia-dawuro-tarcha_zuria"~  "tercha_zuriya",
      adm_1_2_3=="south_west_ethiopia-kefa-goba"~  "goba_sp",
      adm_1_2_3== "oromia-buno_bedele-chewaka" ~ "chwaka",
      adm_1_2_3== "oromia-buno_bedele-chora" ~ "chora_(buno_bedele)",
      adm_1_2_3 == "oromia-buno_bedele-didesa" ~ "dedesa",
      adm_1_2_3 == "oromia-east_hararge-deder_rural" ~ "deder", 
      adm_1_2_3 == "oromia-east_hararge-gursum" ~ "gursum_(or)",
      adm_1_2_3 == "oromia-east_hararge-jarso" ~ "jarso_(east_hararghe)",
      adm_1_2_3 == "oromia-east_hararge-kersa" ~ "kersa_(east_hararge)",
      adm_1_2_3 == "oromia-east_hararge-melka_belo" ~ "melka_balo",
      adm_1_2_3 == "oromia-east_hararge-meyumulke" ~ "meyu_muleke",
      adm_1_2_3 == "oromia-ilu_aba_bora-algie_sachi" ~ "alge_sachi",
      adm_1_2_3 == "oromia-ilu_aba_bora-becho" ~ "becho_(ilu_aba_bora)",
      adm_1_2_3 == "oromia-ilu_aba_bora-bure" ~ "bure_(or)",
      adm_1_2_3 == "oromia-ilu_aba_bora-halu" ~ "halu /huka",
      adm_1_2_3 == "oromia-ilu_aba_bora-hurrumu" ~ "hurumu",
      adm_1_2_3 == "oromia-ilu_aba_bora-mettu" ~ "metu_zuria",
      adm_1_2_3 == "oromia-ilu_aba_bora-mettu_town" ~ "metu_town",
      adm_1_2_3 == "oromia-ilu_aba_bora-yayo" ~ "yayu",
      adm_1_2_3 == "oromia-jimma-agaro" ~ "agaro_town",
      adm_1_2_3 == "oromia-jimma-chora_botor" ~ "chora_(jimma)",
      adm_1_2_3 == "oromia-jimma-gomma" ~ "goma",
      adm_1_2_3 == "oromia-jimma-gumey" ~ "gumay",
      adm_1_2_3 == "oromia-jimma-kersa" ~ "kersa_(jimma)",
      adm_1_2_3 == "oromia-jimma-limu_saka" ~ "limu_seka",
      adm_1_2_3 == "oromia-jimma-mana" ~ "mena_(jimma)",
      adm_1_2_3 == "oromia-jimma-mencho" ~ "mancho",
      adm_1_2_3 == "oromia-jimma-shebe_sembo" ~ "shebe_sambo",
      adm_1_2_3 == "oromia-jimma-sokoru" ~ "sekoru",
      adm_1_2_3 == "oromia-north_shewa_or-girar_jarso" ~ "gerar_jarso",
      adm_1_2_3 == "oromia-north_shewa_or-hidhabu_abote" ~ "hidabu_abote",
      adm_1_2_3 == "oromia-west_hararge-chiro_rural" ~ "chiro_zuria",
      adm_1_2_3 == "oromia-west_hararge-guba_qoricha" ~ "goba_koricha",
      adm_1_2_3 == "oromia-west_hararge-oda_bultum" ~ "kuni_/oda_bultum",
      adm_1_2_3 == "oromia-west_hararge-tulo" ~ "tulo_(or)",
      adm_1_2_3 == "snnp-gofa-gezie_gofa" ~ "gezei_gofa",
      adm_1_2_3 == "snnp-gofa-melokoza" ~ "melekoza",
      adm_1_2_3 == "oromia-west_hararge-burka_dimtu" ~ "burqua_dhintu",
      adm_1_2_3 == "snnp-south_omo-selamago" ~ "salamago",
      adm_1_2_3 == "snnp-south_omo-selamego" ~ "salamago",
      adm_1_2_3 == "snnp-south_omo-semen_ari" ~ "north_ari",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-gide_bench" ~ "gidi_bench",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-guraferda" ~ "gurafereda",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-mizan_aman" ~ "mizan_aman_town",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-shie_bench" ~ "shay_bench",
      adm_1_2_3 == "south_west_ethiopia-bench_sheko-siz_town" ~ "size_town",
      adm_1_2_3 == "south_west_ethiopia-dawuro-esera" ~ "isara",
      adm_1_2_3 == "south_west_ethiopia-dawuro-essera" ~ "isara",
      adm_1_2_3 == "south_west_ethiopia-dawuro-gena_bossa" ~ "gena",
      adm_1_2_3 == "south_west_ethiopia-dawuro-kechi" ~ "kachi",
      adm_1_2_3 == "south_west_ethiopia-dawuro-mermanisa" ~ "mari_mansa",
      adm_1_2_3 == "south_west_ethiopia-dawuro-tercha_city" ~ "tercha_town",
      adm_1_2_3 == "south_west_ethiopia-dawuro-zaba_gazo" ~ "zabagazo",
      adm_1_2_3 == "south_west_ethiopia-kefa-adiyo" ~ "adiyio",
      adm_1_2_3 == "south_west_ethiopia-kefa-bitta" ~ "bita",
      adm_1_2_3 == "south_west_ethiopia-kefa-bonga" ~ "bonga_town",
      adm_1_2_3 == "south_west_ethiopia-kefa-shishoende" ~ "shisho_ande",
      adm_1_2_3 == "south_west_ethiopia-kefa-tello" ~ "tullo",
      adm_1_2_3 == "south_west_ethiopia-mirab_omo-gorigesha" ~ "gori_gesha",
      adm_1_2_3 == "south_west_ethiopia-mirab_omo-meneat_goldia" ~ "menit_goldiye",
      adm_1_2_3 == "south_west_ethiopia-mirab_omo-meneat_shasha" ~ "menit_shasha",
      adm_1_2_3 == "south_west_ethiopia-sheka-andracha" ~ "anderacha",
      adm_1_2_3 == "south_west_ethiopia-sheka-masha_woreda" ~ "masha",
      adm_1_2_3 == "south_west_ethiopia-sheka-teppi_town" ~ "tepi",
      adm_1_2_3 == "south_west_ethiopia-sheka-yeki_woreda" ~ "yeki",
      
      TRUE~ adm3_name
    )
    
  )

check_admin3(df = cleaned_adm3_a,df_master = master_adm)


master_adm_adj |> 
  as_tibble() |> 
  arrange(adm3_en) |>
  print(n=2000) |> 
  filter(
    ### you can filter the concatenated 1_2_3
    str_detect(adm_1_2_3,"gira"),
    
    
    ### if that doesn't work comment it out and
    ### try some pattern matching on other admins
    ###**hint:** "^" = starts with
    ###**        "$" = ends with
   
    # str_detect(adm3_en, "ded")
    # str_detect(adm2_en,"north_shewa_am")
    
    )  |> 
  print(n=1000)


```
