---
title: "Data Quality Visualizations"
author: "Zack Arno"
date: '2022-11-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro



```{r libs_data}
library(tidyverse)
library(here)

invisible(purrr::map(dir(here("R/"),full.names = T),~source(.x,verbose = FALSE,echo = F)))
targets::tar_load_everything()
```


Good no cases where pop treated during current month is > total population
```{r}
pop_treated_gt_total_pop<- RB_pre_post_compiled |> 
  filter(
    popn_treated_during_current_month>total_population
  )
```

We see there are now **`r nrow(pop_treated_gt_total_pop)`** cases where the reported # treated monthly is greater than the total population reported that month.

Let's visually inspect adm2 level monthly reported pop treated vs total population. 

- There are some high outliers? anything to do about these?
- Generally looks pretty decent

```{r}
rb_boxplot(.data = RB_pre_post_compiled,
           x = "adm2_name",
           y = "popn_treated_during_current_month",
           pop = "total_population" )


# log scaled
rb_boxplot(.data = RB_pre_post_compiled,
           x = "adm2_name",
           y = "popn_treated_during_current_month",
           pop = "total_population" ,log = T)



```

I think we should inspect the pop treated over time to make sure the numbers looks good. Alright, I like that, doesn't look lik numbers are getting repeated.

```{r}
RB_pre_post_compiled |> 
  ggplot(aes(x=date,y=popn_treated_during_current_month,color=adm2_name))+
  geom_point()+
  geom_line()+
  scale_y_continuous(labels = scales::comma)+
  scale_x_date(date_breaks = "6 month",date_labels = "%Y-%b",date_minor_breaks = "1 month")+
  theme_bw()+
  labs(y="Population Treated")+
  theme(
    axis.text.x = element_text(angle=90)
  )

```
Let's lookat at `adm3` level post 201905

- hmm that doesn't look too good ... what's going on?
```{r}

RB_post201905_adm3 |> 
  ggplot(
    aes(x= date,
        y=popn_treated_during_current_month,
        color=adm3_name,
        group=adm3_name,
        )
  )+
  geom_path()+
  theme(
    legend.position = "none"
  )

```

```{r}
na_pop_treated_adm3 <- RB_post201905_adm3 |>
  filter(is.na(popn_treated_during_current_month))
```

For `popn_treated_during_current_month`, only **`r nrow(RB_post201905_adm3)`** out of **`nrow(na_pop_treated_adm3)`** values not recorded as `NA`... let's replace those NAs with `0` 

boom, viz makes more sense. Chosen to show each admin 3 as a a line, but color by adm 2 group. Can play with this.
```{r}


RB_post201905_adm3 |> 
  mutate(
  popn_treated_during_current_month=  replace_na(popn_treated_during_current_month,replace = 0)
  ) |> 
    ggplot(
    aes(x= date,
        y=popn_treated_during_current_month,
        color=adm2_name,
        group=adm3_name,
        )
  )+
  geom_path()+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  scale_x_date(date_breaks = "1 month",date_labels = "%Y-%b")+
  labs(y="Pop Treated Current Month")+
  theme_bw()+
  theme(
    axis.text.x = element_text(angle=90),
    legend.position = "none"

  )
  
```


```


Shold fin out a bit more about cumulative all_rounds
```{r}
RB_adm2 |> 
  filter(
    popn_treated_cumulative_all_rounds>total_population
  )

RB_adm2$popn_treated_cumulative_all_rounds
rb_boxplot(.data = RB_adm2,
           x = "adm2_name",
           y = "popn_treated_cumulative_all_rounds",
           pop = "total_population" )
```


Now let's do the same just at the woreda (adm3/district level). We only have woreda level reporting for post 2019-05 data.
```{r}
pop_treated_gt_total_pop_adm3<- RB_post201905_df_compiled |> 
  filter(
    popn_treated_during_current_month>total_population
  )

pop_treated_gt_total_pop_adm3 |> count(date, adm3_name)
```
We see there are now **`r nrow(pop_treated_gt_total_pop_adm3)`** cases where the reported # treated monthly is greater than the total population reported that month.

```{r}
rb_boxplot(.data = RB_post201905_df_compiled,
           x = "adm3_name",
           y = "popn_treated_during_current_month",
           facet = "adm1_name",
           pop = "total_population" )


 # log scaled
rb_boxplot(.data = RB_adm2,
           x = "adm2_name",
           y = "popn_treated_during_current_month",
           pop = "total_population" ,log = T)

```

